---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

1. Load the data (i.e. read.csv())
```{r}
unzip("activity.zip")
activity <- read.csv("activity.csv", header = TRUE)
```

2. Process/transform the data into a format suitable for your analysis - converting the date to the proper format.
```{r}
activity$date <- as.Date(activity$date)
```


## What is mean total number of steps taken per day?

1. The total number of steps taken per day

```{r}
library(plyr)
activity_day <- ddply(activity, .(date), summarize, steps = sum(steps, na.rm = FALSE))
```

2. Histogram of the total number of steps taken each day

```{r}
library(ggplot2)
bw <- diff(range(activity_day$steps, na.rm = TRUE))/7
qplot(steps, data=activity_day, geom="histogram", binwidth = bw)
```

3. The mean and median of the total number of steps taken per day are

```{r}
mean(activity_day$steps, na.rm = TRUE)
median(activity_day$steps, na.rm = TRUE)
```


## What is the average daily activity pattern?

1. Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
activity_interval <- ddply(activity, .(interval), summarize, steps = mean(steps, na.rm = TRUE))
qplot(interval, steps, data = activity_interval, geom = "line")
```

2. The 5-minute interval, on average across all the days in the dataset, which contains the maximun number of steps is `r formatC(activity_interval$interval[which.max(activity_interval$steps)], width=4, format='f', digits=0, flag='0')`.

```{r}
activity_interval$interval[which.max(activity_interval$steps)]
```

## Imputing missing values

1. The total number of missing values in the dataset (i.e. the total number of rows with NAs) is

```{r}
sum(is.na(activity$steps))
```

2. Strategy for filling in all of the missing values in the dataset - for the missing values the mean for the corresponding interval is calculated and the missing value is replaced with it.

```{r}
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
```

3. A new dataset that is equal to the original dataset but with the missing data filled in is created.

```{r}
activity_imp <- ddply(activity, ~ interval, transform, steps = impute.mean(steps))
activity_imp <- activity_imp[order(activity_imp$date, activity_imp$interval), ]
```
4. Histogram of the total number of steps taken each day. 

```{r}
activity_day <- ddply(activity_imp, .(date), summarize, steps = sum(steps))
qplot(steps, data=activity_day, geom="histogram", binwidth = bw)
```

The mean and median total number of steps taken per day

```{r}
mean(activity_day$steps)
median(activity_day$steps)
```

The value of the mean remains exactky the same (due to the method of imputing missing values I chose), the median changes slightly.

## Are there differences in activity patterns between weekdays and weekends?

1. A new factor variable is created in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r}
activity_imp$wd <- as.factor(ifelse((weekdays(activity_imp$date) == "Saturday" | weekdays(activity_imp$date) == "Sunday"), "weekend", "weekday"))
```

2. Panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

```{r}
activity_di <- ddply(activity_imp, .(interval, wd), summarize, steps = mean(steps, na.rm = TRUE))
ggplot(data = activity_di, aes(interval, steps)) + geom_line() + facet_grid(wd ~ .)
```
